Synthesising Complex Termination Arguments

### 2014-06-23 papers/termination/paper.tex
2nd order termination tweak
2nd-order termination

The existence of a ranking function for a single loop is equivalent to the satisfiability
of the formula from Definition~\ref{def:termination-formula}.  If this formula is satisfiable,
$R$ is a ranking function that proves the loop $L$ terminates, so the termination of $L$
is equivalent to the satisfiability of Definition~\ref{def:termination-formula}.

@@

The conditional termination formula becomes more complex in the presence of multiple loops.
Consider the following program with two loops $L_1$ and $L_2$:

@@

We can specify the existence of a ranking function, and therefore the termination
of a program, with the second order formulae defined in Fig.~\ref{fig:termination-formulae}.
We can observe that each of these formulae (with the exception of Definition~\ref{def:nonterm-formula})
is an $\exists \forall$ formula where the the universally quantified variables only range over
ground terms.
The program synthesis problem can be described as finding a satisfying assigment to the
synthesis formula:

>>>>

The existence of a ranking function for a single loop is equivalent to the satisfiability
of the formula from Definition~\ref{def:termination-formula}.  If this formula is satisfiable,
$R$ is a ranking function that proves the loop $L$ terminates.

@@

The conditional termination formula becomes more complex in the presence of multiple loops.
Consider the following program with two loops $L_1$ and $L_2$, one after the other:

@@

With the exception of Definition~\ref{def:nonterm-formula}, each of the second-order formulae
above is an $\exists \forall$ formula where the the universally quantified variables only range over
ground terms.  We refer to this syntactic class as the \emph{synthesis fragment} of second-order
logic, because program synthesis problem can be described as finding a satisfying assigment to the
following formula:
